### 2014-10-22 doc/thesis/evaluation.tex
Typing the core of luasocket
First version of the luasocket description files
Fixing evaluation to include methods from userdata
Improving the text about subtyping rules
Reorganizing the chapter about evaluation and writing the section about the MD5 case study

We performed some case studies on existing Lua libraries
to evaluate the design of our type system.
For each library, we used Typed Lua to either rewrite its modules
or to write statically typed interfaces to its modules through
Typed Lua's description files.
In this chapter we present our evaluation results and discuss
some interesting cases.

Tables \ref{tab:evalbylib} and \ref{tab:evalbymod} sumarizes our
evaluation results:
the first presents the results agruped by library while
the second presents the results agruped by module.
We split the members that we typed into three categories:
\emph{easy}, \emph{poly}, and \emph{hard}.
The \emph{easy} category represents the members that we could give
a precise static type for them.
The \emph{poly} category represents the members that we had to
rely on the dynamic type, and we believe we could give a precise
static type for them if our type system had parametric polymorphism.
The \emph{hard} category represents the members that we could not
give a precise static type for them, relying on the dynamic type
in most of the cases, but with no hopes that parametric polymorphism
would help us to give a precise static type for them.
The last column of each table contains the actual evaluation result
of each case study, that is basically the percentage of members
that are under the \emph{easy} category.

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
Lua Standard Libraries & 87 & 11 & 25 & 123 & 71\% \\
\hline
MD5 & 7 & 0 & 0 & 7 & 100\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation of Typed Lua by library}
\label{tab:evalbylib}
\end{table}

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
\multirow{9}{*}{Lua Standard Libraries}
& base & 9 & 5 & 12 & 26 & 35\% \\
\cline{2-7}
& coroutine & 0 & 0 & 6 & 6 & 0\% \\
\cline{2-7}
& package & 5 & 2 & 1 & 8 & 62\% \\
\cline{2-7}
& string & 13 & 0 & 1 & 14 & 93\% \\
\cline{2-7}
& table & 1 & 4 & 1 & 6 & 17\% \\
\cline{2-7}
& math & 28 & 0 & 1 & 29 & 97\% \\
\cline{2-7}
& bit32 & 12 & 0 & 0 & 12 & 100\% \\
\cline{2-7}
& io & 9 & 0 & 2 & 11 & 82\% \\
\cline{2-7}
& os & 10 & 0 & 1 & 11 & 91\% \\
\hline
\multirow{1}{*}{MD5}
& md5 & 7 & 0 & 0 & 7 & 100\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation of Typed Lua by module}
\label{tab:evalbymod}
\end{table}

>>>>

We performed some case studies on existing Lua libraries
to evaluate the design of our type system.
For each library, we used Typed Lua to either rewrite its modules
or to write statically typed interfaces to its modules through
Typed Lua's description files.
In this chapter we present our evaluation results, discuss
some interesting cases, and compare our type system to
related work.

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
Lua Standard Libraries & 92 & 11 & 27 & 130 & 71\% \\
\hline
MD5 & 13 & 0 & 0 & 13 & 100\% \\
\hline
Lua Socket & 85 & 22 & 19 & 126 & 67\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for each case study}
\label{tab:evalbycase}
\end{table}

Table \ref{tab:evalbycase} sumarizes our evaluation results for each
case study that we used Typed Lua for typing their members.
We split the members that we typed into three categories:
\emph{easy}, \emph{poly}, and \emph{hard}.
The \emph{easy} category represents the members that we could give
a precise static type for them.
The \emph{poly} category represents the members that we believe that
parametric polymorphism would help give a precise static type for them,
and we had to rely on the dynamic type to type them because our
type system does not support parametric polymorphism.
The \emph{hard} category represents the members that even parametric
polymorphism would not help give a precise static type for them,
and we also had to rely on the dynamic type to type them;
for instance, this category includes functions that require
intersection types to describe their precise static type.
The last column of the table shows the percentage of members that
are under the \emph{easy} category for each case study.
This percentage is our evaluation of Typed Lua, as it
represents how much static typing we could introduce to each one of
our case studies.

Before comparing our type system to related work, we will discuss
each case study in more detail.
For each case study, we will split the evaluation results according
to the modules that each one of them include.
This shall allow us to better discuss the contributions and limitations
of our type system.

\section{Lua Standard Libraries}

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
\multirow{9}{*}{Lua Standard Libraries}
& base & 9 & 5 & 12 & 26 & 35\% \\
\cline{2-7}
& coroutine & 0 & 0 & 6 & 6 & 0\% \\
\cline{2-7}
& package & 5 & 2 & 1 & 8 & 62\% \\
\cline{2-7}
& string & 13 & 0 & 1 & 14 & 93\% \\
\cline{2-7}
& table & 1 & 4 & 1 & 6 & 17\% \\
\cline{2-7}
& math & 28 & 0 & 1 & 29 & 97\% \\
\cline{2-7}
& bit32 & 12 & 0 & 0 & 12 & 100\% \\
\cline{2-7}
& io & 14 & 0 & 4 & 18 & 78\% \\
\cline{2-7}
& os & 10 & 0 & 1 & 11 & 91\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for Lua Standard Libraries}
\label{tab:evallsl}
\end{table}

\section{MD5}

The MD5 library \citep{lmd5} is the case study that we used to
introduce Typed Lua's description files.
We chose this case study because it is simple and contains just
one module, being a good example for introducing how we can use
Typed Lua's description files for giving a statically typed
interface to a Lua library that is written in C, and to 
introduce how we can type an \emph{userdata}.
Table \ref{tab:evalmd5} sumarizes the evaluation results for MD5.

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
\multirow{1}{*}{MD5}
& md5 & 13 & 0 & 0 & 13 & 100\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for MD5}
\label{tab:evalmd5}
\end{table}

Even tough the MD5 library is quite simple and it was quite easy to type,
we found a little discrepancy between its documentation and its static
typing.
The documentation suggested that the typing of the function \texttt{update}
should be \texttt{(md5\string_context, string) -> (md5\string_context)},
and a comment in the source code suggested that it should be
should be \texttt{(md5\string_context, string, string*) -> (md5\string_context)}.
However, while testing it and reading its source code, we noticed that
its actual type is \texttt{(md5\string_context, string*) -> (md5\string_context)},
that is, we can pass zero or more strings to \texttt{update}.

This case study shows a connection between documentation and software testing.
Even though type annotations are a good source of documentation, this case
study shows that identifying what functions discard extra parameters can be
not so obvious for someone that is typing an external library.
Reading the source code and the library's test script was essential to confirm
the typing of \texttt{update}.

\section{Lua Socket}

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
\multirow{7}{*}{Lua Socket}
& socket & 50 & 3 & 7 & 60 & 83\% \\
\cline{2-7}
& ftp & 6 & 1 & 1 & 8 & 75\% \\
\cline{2-7}
& http & 3 & 0 & 2 & 5 & 60\% \\
\cline{2-7}
& ltn12 & 10 & 10 & 0 & 20 & 50\% \\
\cline{2-7}
& mime & 2 & 8 & 7 & 17 & 12\% \\
\cline{2-7}
& smtp & 6 & 0 & 2 & 8 & 75\% \\
\cline{2-7}
& url & 8 & 0 & 0 & 8 & 100\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for Lua Socket}
\label{tab:evalsocket}
\end{table}

\section{Related Work}
